---
title: Graphing Stuff
subtitle: With SPSS
author: Stats Central
institute: UNSW
fontsize: 11pt
params: 
  R:  false
  SPSS: true 
# output:
#   revealjs::revealjs_presentation:
#     theme: sky
#     highlight: pygments
#     center: true
#     transition: none
output:
  beamer_presentation:
   includes:
    in_header: unsw_beamer.tex
   keep_tex: true
   latex_engine: pdflatex
   theme: Madrid
   colortheme: seahorse
   slide_level: 3
make149: true
classoption:
- twocolumn
---

```{r setup, include = FALSE, fig.align = "center"}

require(knitr)
knitr::opts_chunk$set(echo = TRUE, fig.width=9, fig.height = 5, fig.align = "center")

```

```{r include = F, message = F, warning = F}

#install.packages("ggpubr", dependencies = TRUE)

library(MindOnStats)
library(RcmdrMisc)
library(ggplot2)
library(ggpubr)
library(tidyverse)
library(boot)
library(kableExtra)
library(ggplot2)
```


### What am I talking about? {.fragile} 

- Making good graphics for publication
- Basic principles of graphing
- Building blocks of graphs
- Examples of common types of graph


### Statistical Graphs

“… statistical graphics are instruments to help people reason about quantitative information”^[Tufte, E. R. 2001. 'The Visual Display of Quantitative Information', p. 91]

- Should be illustrative
- Should be helpful

Tell your story


### Basic elements of a plot

- Graph panel
- Scales, labels, tick marks
- Plotting symbols, line types, colour
- Reference lines
- Legend
- Captions


Maximise the “data-ink ratio”^[Tufte, E. R. 2001. 'The Visual Display of Quantitative Information', p. 96]

- Use ink for data, not unnecessary decoration


### Basic guidelines of graphing data

**Make the data stand out**

- Let the data fill the graph panel
- Don’t force axes to start at zero if data values are not close to zero (could use “break” marks to show axis does not start at zero)
- Label the axes clearly
- Use symbols, lines, colours etc. that highlight the data
- Use reference lines or grids with care
- Put legends outside the graph panel
- Put notes in the caption or text
- Make sure the graph remains clear if made smaller (e.g. for publication)

### Colour

Colour - symbols and lines

some people are colourblind
hard to see differences especially among reds, oranges, yellows and greens
there are different types of colourblindness
don’t use red and green together
check out ColorBrewer for colour palettes (in R, RColorBrewer package)

Avoid “chartjunk”^[Tufte, E. R. 2001. 'The Visual Display of Quantitative Information', p. 104] - more about this later

###

Make sure the graph is properly understood

Put all necessary details about the graph in the caption
describe everything that is graphed
clearly explain any error bars
note important features in the data
describe the conclusions that are made from the data shown

### Variation

Two areas for showing variation:

- Empirical variation of data - exploring data
- Variation of statistics - exploring data and model output

### Empirical variation - discrete data

Data: A sample of three types of gastropod shell was scored as occupied by a hermit crab or empty5.

Question: Do hermit crabs prefer a certain shell type?

```{r}
# create optometrist N data
opt1 <- data.frame(Occupation=c("Optometrist", "Optometrist", "Optometrist", "Optometrist", "Optometrist","Ophthamologist","Ophthamologist","Ophthamologist","Ophthamologist","Ophthamologist"),
                 Location=c("Major City", "Inner Regional","Outer Regional","Remote","Rural","Major City", "Inner Regional","Outer Regional","Remote","Rural"),
                 N=c(4221,	811,	252,	23,	17,
                     818,  121,	23,	NA,   NA),
                 FTE=c(3796.5,   765.3,	242.5,	24.0,	16.9,
                       850.7,	127.1,	27.0,	NA,   NA),
                 FTErate=c(20.7,	17.0,	11.8,	8.3,	8.4,
                       4.6,	2.8,	1.3,	NA,   NA))

opttable<-opt1 %>%
   dplyr::select(Occupation, Location,N) %>%
   rename(value=N) %>%
   pivot_wider(values_from = value, names_from = Location)
opt2<-opt1 %>%
   dplyr::select(Occupation, Location, FTE) %>%
   rename(value=FTE) %>%
   pivot_wider(values_from = value, names_from = Location)
opt3<-opt1 %>%
   dplyr::select(Occupation, Location, FTErate) %>%
   rename(value=FTErate) %>%
   pivot_wider(values_from = value, names_from = Location)
opttable <- rbind(opttable,opt2, opt3)
rm(opt2,opt3)
opttable%>% 
   kbl() %>%
  kable_classic(full_width = F) %>%
  pack_rows("N", 1, 2) %>%
  pack_rows("FTE", 3, 4) %>%
  pack_rows("FTE per 100,000", 5, 6) 


# bar plot for Ns. Not so great because we can't 

#

# Table 2.1: Full-time equivalent (FTE) number and rate of eye-care providers, by remoteness area, 2019 Source: AIHW analysis of National Health Workforce Dataset (DOH 2020b)
					

```

